Tracing the Hercules stream around the Galaxy

### 2010-09-12 tex/response.txt

	 The paper titled Tracing the Hercules stream around the Galaxy by Jo
	 Bovy is concerned with the origin and detectability of this
	 elusive stream. Overall the paper is well written and
	 clear. I have a few major concerns and some minor concerns
	 that should be addressed before the paper can be accepted for
	 publication.

	 Major concerns:

	 1. Sect 4, page 6, last paragraph: The author discusses in
	 some detail, and shows plots to illustrate, the possible
	 implications of e.g.  errors in the distance determination
	 etc. The one error that is not address explicitly in the text
	 or plots is the error on the measured radial velocities - how
	 well do we need to determine these velocities in order to be
	 able to unambiguously detect the Hercules stream? 1, 5, or 10
	 km/s? Or some other number? And what happens when V0 is
	 absolutely known or if it also has an error? Perhaps the
	 result is that the errors do no matter, but in that case this
	 should be shown and arguments to why it does not matter
	 should be given. Also, will either the RVS on board Gaia or
	 APOGEE deliver radial velocities with small enough errors to
	 detect the Hercules stream?

Gaia errors: see below; APOGEE errors: better than 0.5 km/s.

@@

	 2. A brief discussion about which tracers, e.g. dwarf stars,
	 red clump, red giant stars, to be used to test the model
	 should also be included. For example, is it necessary to
	 reach as far away from the solar neighbourhood as possible
	 (eg giants), or is the major test to get as high volume
	 density of tracers as possible (eg cool dwarfs). Such a
	 discussion would be highly useful for anyone wanting to
	 construct surveys for the very purpose of detecting
	 Hercules. This discussion does not need to be long.

> 3 kpc: giants (dwarfs in GAIA few kpc, giants go far enough and
  bright enough to have RVs)

@@

	 3. I find Discussion in Sect 5 (for both Gaia and APOGEE)
	 somewhat sketchy. For example, no mention of the fact that
	 Gaia will provide radial velocities for stars down to roughly
	 17 magnitude is mentioned.  Should the reader conclude that
	 Hercules could be detected with the tangential velocities
	 alone? In that case, how does that work with the tests
	 presented in this article? Again, here the intrinsic
	 brightness of the tracers is of some interest.

I added a reference to the line-of-sight velocity accuracies expected
for Gaia, and an explicit comment that the velocities of giants will
be measured accurately enough to verify the predictions of this paper.

### 2010-09-12 tex/hercules.tex
groups-simulations: add HERMES reference
groups-simulations: bar olr figure
groups-simulations: further Gaia discussion

The astrometric \Gaia\ mission will measure the parallaxes and proper
motions of up to one billion stars, most of which will be disk stars
\citep{Perryman01a}. With parallax and proper motion accuracies down
to 10 microarcseconds, \Gaia\ will be able to probe the kinematics of
the disk out to several kpc in all directions
\citep{bailerjones08a}. Crucially, it will provide large samples of
stars in many of the promising regions in
\figurename~\ref{fig:rphi2d}, which will be able to verify the
predictions laid out in this paper. This will also allow the dynamical
properties of the bar to be tightly constrained, something that will
be hard for \Gaia\ to do using direct observations of the bulge due to
the large extinction toward the Galactic center \citep{robin05a}.

@@

\acknowledgements It is a pleasure to thank the anonymous referee,
Mike Blanton, and David W.~Hogg for helpful discussions and
assistance.  Financial support for this project was provided by the
National Aeronautics and Space Administration (grant NNX08AJ48G) and
the National Science Foundation (grant AST-0908357).

@@

@@

@@

\clearpage
\begin{figure}
\includegraphics[width=0.35\textwidth]{distuncertain.ps}
\includegraphics[width=0.35\textwidth]{vlosuncertain.ps}\\
\includegraphics[width=0.35\textwidth]{df.ps}
\includegraphics[width=0.35\textwidth]{slope.ps}\\
\includegraphics[width=0.35\textwidth]{barstrength.ps}
\caption{Variation of the predicted velocity distribution at R = 1.075
  \Ro , $\phi = -21^{\circ}$ (d= 0.4 \Ro, $l$ = 270$^{\circ}$) with
  distance uncertainties; line-of-sight velocity uncertainties;
  parameters and shape of the distribution function; shape of the
  rotation curve; and bar strength.}\label{fig:1dvar}
\end{figure}

>>>>

The astrometric \Gaia\ mission will measure the parallaxes and proper
motions of up to one billion stars, most of which will be disk stars
\citep{Perryman01a}. With parallax and proper motion accuracies down
to 10 microarcseconds and mission-averaged line-of-sight velocity
uncertainties smaller than 10 km s$^{-1}$ for many stars to 17th
magnitude, \Gaia\ will be able to probe the kinematics of the disk out
to several kpc in all directions
\citep{Katz04a,bailerjones08a}. Crucially, it will provide large
samples of giants in many of the promising regions in
\figurename~\ref{fig:rphi2d}, with velocity measurements accurate
enough to verify the predictions laid out in this paper. This will
also allow the dynamical properties of the bar to be tightly
constrained, something that will be hard for \Gaia\ to do using direct
observations of the bulge due to the large extinction toward the
Galactic center \citep{robin05a}.

@@

\citep{Freeman10a}.

\acknowledgements It is a pleasure to thank the anonymous referee for
valuable comments and Mike Blanton and David W.~Hogg for helpful
discussions and assistance.  Financial support for this project was
provided by the National Aeronautics and Space Administration (grant
NNX08AJ48G) and the National Science Foundation (grant AST-0908357).

@@

\bibitem[Freeman \etal(2010)]{Freeman10a}
  Freeman,~K., Bland-Hawthorn,~J., \& Barden,~S., 2010, 
  AAO Newsletter, 117, 9

@@

\bibitem[Katz \etal(2004)]{Katz04a}
Katz,~D., \etal, 2004,
\mnras, 354, 1223

@@

\clearpage
\begin{figure}
\includegraphics[width=0.35\textwidth]{distuncertain.ps}
\includegraphics[width=0.35\textwidth]{vlosuncertain.ps}\\
\includegraphics[width=0.35\textwidth]{df.ps}
\includegraphics[width=0.35\textwidth]{slope.ps}\\
\includegraphics[width=0.35\textwidth]{rolr.ps}
\includegraphics[width=0.35\textwidth]{barstrength.ps}
\caption{Variation of the predicted velocity distribution at R = 1.075
  \Ro , $\phi = -21^{\circ}$ (d= 0.4 \Ro, $l$ = 270$^{\circ}$) with
  distance uncertainties; line-of-sight velocity uncertainties;
  parameters and shape of the distribution function; shape of the
  rotation curve; and bar strength.}\label{fig:1dvar}
\end{figure}
